---
title: "Covid GeoStats"
output: 
  flexdashboard::flex_dashboard:
    theme: cosmo
runtime: shiny
---

```{r setup, include=FALSE}
library(rsconnect)
library(flexdashboard)
library(DT)
library(plotly)
library(RColorBrewer)

library(leaflet) 
library(leaflet.extras)

library(sf)
library(rmapshaper)
#download summary data 
PHU_Summary <- read.csv("./dataset/PHU_Summary.csv", encoding = "UTF-8")
#read spatial boundaries data 
PHU_bounds <- readRDS("./spatial files/PHU_Boundaries.rds")

# simplify the boundary file
# this will help the maps run so much faster when rendering 
PHU_bounds <- rmapshaper::ms_simplify(PHU_bounds, keep = 0.01, keep_shapes = TRUE)
```


GIS
===================================== 

Column {data-width=600}
-------------------------------------
    
### Map
    
```{r}
csum_map <- merge(PHU_bounds, PHU_Summary, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )
pal <- colorNumeric("viridis", domain = csum_map$Total_Cases, reverse = F)

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Total_Cases == max(Total_Cases))) %>% 
  as.vector()

csum_map$popup_text <- 
  paste0('<h4>', csum_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Total Cases: ', '<b>', format(csum_map$Total_Cases, big.mark = ","),'</b>','</br>',
         sep = " ") %>% 
  
  lapply(htmltools::HTML)


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addProviderTiles(providers$CartoDB.Positron) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Total_Cases), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T),
              
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              popup = ~popup_text) %>%
  
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  addLegend("bottomright", pal = pal, values = csum_map$Total_Cases, 
            title = "Total Cases", 
            opacity = 0.5)
```
   
Column {data-width=400}
-------------------------------------
   
### Table

```{r}
DT::datatable(data = PHU_Summary[, c("Reporting_PHU", "Total_Cases")] %>% dplyr::arrange(desc(Total_Cases)),
              class = 'cell-border stripe',
              colnames = c("Public Health Unit", "Total Cases"), 
              rownames = F, 
              options = list(dom = 't',ordering = F, 
                             autoWidth = T)
)
```   
 
### Histogram
    
```{r}
PHU_Summary$Reporting_PHU <- factor(PHU_Summary$Reporting_PHU, 
                                      levels = unique(PHU_Summary$Reporting_PHU)[order(PHU_Summary$Total_Cases, decreasing = F)])

plot_ly(data = PHU_Summary, 
        y = ~Total_Cases, 
        x = ~Reporting_PHU, 
        type = 'bar', 
        name = "Total Cases",
        text = ~paste(Reporting_PHU, 
                      "<br>", 
                      format(PHU_Summary$Total_Cases, nsmall = 0, big.mark = ","), 
                      "Total Cases", 
                      sep = " "), 
        
        hoverinfo = "text", 
        marker = list(color = '#6EC5E9')
        ) %>% 
  
  layout(title = "<b> Total Cases Per Health Region </b>",
         xaxis = list(title = "", zeroline = FALSE, showline = FALSE, 
                      showgrid = FALSE, showticklabels = FALSE), 
         yaxis = list(title = "Total Cases"), 
         barmode = 'stack',
         hovermode = "y-unified", 
         hoverlabel=list(bgcolor="black"))
```


Per Capita Rates
===================================== 

Column {data-width=600 .tabset .tabset-fade}
-------------------------------------
    
### Cases

```{r}
pal <- colorNumeric("viridis", domain = csum_map$Cases_pC, reverse = F)

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Cases_pC == max(Cases_pC))) %>% 
  as.vector()

csum_map$popup_text <- 
  paste0('<h4>', csum_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Cases per Capita: ', '<b>', format(csum_map$Cases_pC, big.mark = ","),'</b>','</br>',
         sep = " ") %>% 
  
  lapply(htmltools::HTML)


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addProviderTiles(providers$CartoDB.Positron) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Cases_pC), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T),
              
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              popup = ~popup_text) %>%
  
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  addLegend("bottomright", pal = pal, values = csum_map$Cases_pC, 
            title = "Cases per Capita (100,000)", 
            opacity = 0.5)
```

### Recovered 

```{r}
pal <- colorNumeric("viridis", domain = csum_map$Recovered_pC, reverse = F)

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Recovered_pC == max(Recovered_pC))) %>% 
  as.vector()

csum_map$popup_text <- 
  paste0('<h4>', csum_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Recoveries per Capita: ', '<b>', format(csum_map$Recovered_pC, big.mark = ","),'</b>','</br>',
         sep = " ") %>% 
  
  lapply(htmltools::HTML)


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addProviderTiles(providers$CartoDB.Positron) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Recovered_pC), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T),
              
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              popup = ~popup_text) %>%
  
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  addLegend("bottomright", pal = pal, values = csum_map$Recovered_pC, 
            title = "Recoveries per Capita (100,000)", 
            opacity = 0.5)
```

### Actives 

```{r}
pal <- colorNumeric("viridis", domain = csum_map$Actives_pC, reverse = F)

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Actives_pC == max(Actives_pC))) %>% 
  as.vector()

csum_map$popup_text <- 
  paste0('<h4>', csum_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Active Cases per Capita: ', '<b>', format(csum_map$Actives_pC, big.mark = ","),'</b>','</br>',
         sep = " ") %>% 
  
  lapply(htmltools::HTML)


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addProviderTiles(providers$CartoDB.Positron) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Actives_pC), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T),
              
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              popup = ~popup_text) %>%
  
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  addLegend("bottomright", pal = pal, values = csum_map$Actives_pC, 
            title = "Active Cases per Capita (100,000)", 
            opacity = 0.5)
```

### Deaths 

```{r}
pal <- colorNumeric("viridis", domain = csum_map$Deaths_pC, reverse = F)

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Deaths_pC == max(Deaths_pC))) %>% 
  as.vector()

csum_map$popup_text <- 
  paste0('<h4>', csum_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Deaths Cases per Capita: ', '<b>', format(csum_map$Deaths_pC, big.mark = ","),'</b>','</br>',
         sep = " ") %>% 
  
  lapply(htmltools::HTML)


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addProviderTiles(providers$CartoDB.Positron) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Deaths_pC), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T),
              
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              popup = ~popup_text) %>%
  
  addFullscreenControl() %>% 
  addResetMapButton() %>% 
  addLegend("bottomright", pal = pal, values = csum_map$Deaths_pC, 
            title = "Deaths per Capita (100,000)", 
            opacity = 0.5)
```

### Close Contact 
### Outbreak
### Travel 
### Unknown 

Column {data-width=400}
-------------------------------------
   
### Table

### Histogram

Acquisition Type
===================================== 

Column {data-width=600}
-------------------------------------
    
### Dominant Acquisiton Type
    
```{r}
#finding the most common acquistion method 
PHU_Summary$AcqMax <- colnames(PHU_Summary[,c("CloseContact", "Outbreak", "Travel", "Unknown")])[apply(PHU_Summary[,c("CloseContact", "Outbreak", "Travel", "Unknown")],1,which.max)]

Acq_map <- merge(PHU_bounds, PHU_Summary, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )

Acq_map$AcqMax <- ifelse(Acq_map$AcqMax == "CloseContact", paste0("Close Contact"), Acq_map$AcqMax)

#factpal <- colorFactor(topo.colors(3), Acq_map$AcqMax)
AcqCats <- colorFactor(palette = c("#6EC5E9", "#F4B400", "#9b4a11"), 
                       levels = c("Close Contact", "Outbreak", "Unkown"),
                       na.color = "#808080")

Acq_map$popup_text <- 
  paste0('<h4>', Acq_map$PHU_NAME_ENG, '</h4>', '<br>',
         'Close Contact: ', '<b>', format(Acq_map$CloseContact, big.mark = ","), ' (', Acq_map$CloseContact_Pct, '%) ','</b>','</br>',
         'Outbreak: ', '<b>', format(Acq_map$Outbreak, big.mark = ","), ' (', Acq_map$Outbreak_Pct, '%) ',  '</b>', '</br>',
         'Travel: ', '<b>', format(Acq_map$Travel, big.mark = ","), ' (', Acq_map$Travel_Pct, '%) ',  '</b>', '</br>',
         'Unknown: ', '<b>', format(Acq_map$Unknown, big.mark = ","), ' (', Acq_map$Unknown_Pct, '%) ',  '</b>', 
         sep = " ") %>% lapply(htmltools::HTML)

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  # setView(lat = mean(bbox[2], bbox[4]), 
  #         lng = mean(bbox[1], bbox[3]), 
  #         zoom = 9) %>%
  
  addPolygons(data  = Acq_map,
              fillColor = ~AcqCats(AcqMax), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T), 
              label =  ~as.character(PHU_NAME_ENG),
              labelOptions = labelOptions(noHide = F,direction = "top", textsize = "15px"),
              
              popup = ~popup_text)%>%
  
  addFullscreenControl() %>%
  addResetMapButton() %>% 
  addLegend(pal = AcqCats, 
            values = Acq_map$AcqMax,
            title = NULL, 
            opacity = 0.7, 
            position  = "bottomright")
```
   
Column {data-width=400}
-------------------------------------
   
### Table

```{r}
DT::datatable(data = PHU_Summary[, c("Reporting_PHU", "CloseContact", "Outbreak", "Travel", "Unknown")] %>% dplyr::arrange(desc(CloseContact)),
              class = 'cell-border stripe',
              colnames = c("Public Health Unit", "Close Contact", "Outbreak", "Travel", "Unknown"), 
              rownames = F, 
              options = list(dom = 't',ordering = F, page_length = 34, lengthChange = FALSE,
                             columnDefs = list(list(className = 'dt-center', targets="_all")))
)
```   
 
### Histogram
    
```{r}
```



Hotspot Analysis
===================================== 

Column {data-width=600}
-------------------------------------
    
### Map

Column {data-width=400}
-------------------------------------
   
### Table

### Histogram


About this Site
===================================== 

#### Last Update  

`r format(Sys.time(), "%d %B, %Y")`

#### Background
*"Nothing in life is to be feared, it is only to be understood. Now is the time to understand more, so that we may fear less" - Marie Curie*  


#### Code
Code and input data used to generate this web app are available on [Github](https://github.com/Erebus54/covid-geostats).

#### Sources 

**Status of COVID-19 cases in Ontario** : [Open Data Ontario](https://data.ontario.ca/dataset/status-of-covid-19-cases-in-ontario), updated daily, which presents a breakdown by region, case status, age range of cases 

**Census Profile, 2016 Census** : [StatsCan](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/index.cfm?Lang=E&TABID=1), queried this to obtain the areal size and population figures to compute the per capita metrics 

**Ministry of Health Public Health Unit Boundaries** : [Ontario GeoHub](https://geohub.lio.gov.on.ca/datasets/d8fba69152e4408dabfe70e85a2688d2_44), contains the geographic boundaries of all 34 Public Health Units (PHUs) in Ontario, based on Statistics Canada’s Health Region Boundary File (2018)

#### Author 

Patrick Schnurbusch 

#### Contact 

patrick.schnurbusch@gmail.com
