---
title: "Covid GeoStats"
output: 
  flexdashboard::flex_dashboard:
    #orientation: rows
    social: menu
    theme: paper
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(rsconnect)
library(flexdashboard)
library(DT)
library(plotly)
library(RColorBrewer)

library(leaflet) 
library(leaflet.extras)

library(sf)

#download summary data 
PHU_Summary <- read.csv("./dataset/PHU_Summary.csv", encoding = "UTF-8")
#read spatial boundaries data 
PHU_bounds <- readRDS("./spatial files/PHU_Boundaries.rds")
```


GIS
===================================== 

Column {data-width=600}
-------------------------------------
    
### Map
    
```{r}

csum_map <- merge(PHU_bounds, PHU_Summary, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )
pal <- colorNumeric("viridis", domain = csum_map$Total_Cases, reverse = F)

popup = paste0("<b>Country: </b>")

#bbox <- st_bbox(csum_map) %>% 
 # as.vector()

#centers focus to current highest region via total cases 
bbox <- st_bbox(csum_map %>% dplyr::filter(Total_Cases == max(Total_Cases))) %>% 
  as.vector()


leaflet() %>% 
  
  #fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%

  #addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  
  addPolygons(data  = csum_map,
              fillColor = ~pal(Total_Cases), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 0.5, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 0.5, 
                bringToFront = T), 
              
              # labelOptions = labelOptions(
              #   style = list("font-weight" = "normal", padding = "3px 8px"),
              #   textsize = "15px",
              #   direction = "auto")
              ) %>%
  
  addFullscreenControl() %>% 
  
  addResetMapButton() %>% 
  
  addLegend("bottomright", pal = pal, values = csum_map$Total_Cases, 
            title = "Total Cases", 
            opacity = 0.5)
```
   
Column {data-width=400}
-------------------------------------
   
### Table

```{r}
DT::datatable(data = PHU_Summary[, c("Reporting_PHU", "Total_Cases")] %>% dplyr::arrange(desc(Total_Cases)),
              class = 'cell-border stripe',
              colnames = c("Public Health Unit", "Total Cases"), 
              rownames = F, 
              options = list(dom = 't',ordering = F, page_length = 34, lengthChange = FALSE,
                             columnDefs = list(list(className = 'dt-center', targets="_all")))
)
```   
 
### Histogram
    
```{r}
PHU_Summary$Reporting_PHU <- factor(PHU_Summary$Reporting_PHU, 
                                      levels = unique(PHU_Summary$Reporting_PHU)[order(PHU_Summary$Total_Cases, decreasing = F)])

plot_ly(data = PHU_Summary, 
        y = ~Total_Cases, 
        x = ~Reporting_PHU, 
        type = 'bar', 
        name = "Total Cases",
        text = ~paste(Reporting_PHU, 
                      "<br>", 
                      format(PHU_Summary$Total_Cases, nsmall = 0, big.mark = ","), 
                      "Total Cases", 
                      sep = " "), 
        
        hoverinfo = "text", 
        marker = list(color = '#6EC5E9')
        ) %>% 
  
  layout(title = "<b> Total Cases Per Health Region </b>",
         xaxis = list(title = "", zeroline = FALSE, showline = FALSE, 
                      showgrid = FALSE, showticklabels = FALSE), 
         yaxis = list(title = "Total Cases"), 
         barmode = 'stack',
         hovermode = "y-unified", 
         hoverlabel=list(bgcolor="black"))
```


Outcomes 
===================================== 

Acquisition Type
===================================== 

Column {data-width=600}
-------------------------------------
    
### Dominant Acquisiton Type via Cases 
    
```{r}

#finding the most common acquistion method 
PHU_Summary$AcqMax <- colnames(PHU_Summary[,c("CloseContact", "Outbreak", "Travel", "Unknown")])[apply(PHU_Summary[,c("CloseContact", "Outbreak", "Travel", "Unknown")],1,which.max)]

Acq_map <- merge(PHU_bounds, PHU_Summary, by.x = "PHU_NAME_ENG", by.y = "Reporting_PHU" )

Acq_map$AcqMax <- ifelse(Acq_map$AcqMax == "CloseContact", paste0("Close Contact"), Acq_map$AcqMax)

#factpal <- colorFactor(topo.colors(3), Acq_map$AcqMax)
AcqCats <- colorFactor(palette = c("#6EC5E9", "#F4B400", "#9b4a11"), 
                       levels = c("Close Contact", "Outbreak", "Unkown"))
leaflet() %>% 
  
  setView(lat = mean(bbox[2], bbox[4]), 
          lng = mean(bbox[1], bbox[3]), 
          zoom = 9) %>%
  
  addPolygons(data  = Acq_map,
              fillColor = ~AcqCats(AcqMax), 
              weight = 1, 
              opacity = 1, 
              
              color = "white", 
              dashArray = 3, 
              fillOpacity = 1, 
              
              highlightOptions = highlightOptions(
                weight = 2, 
                color = "red", 
                fillOpacity = 1, 
                bringToFront = T), 
              
              # labelOptions = labelOptions(
              #   style = list("font-weight" = "normal", padding = "3px 8px"),
              #   textsize = "15px",
              #   direction = "auto")
  ) %>%
  
  addFullscreenControl() %>% 
  
  addResetMapButton() %>% 
  
  addLegend("bottomright", pal = AcqCats, values = Acq_map$AcqMax, 
            title = "Acquisition Type", 
            opacity = 1)
```
   
Column {data-width=400}
-------------------------------------
   
### Table

```{r}
```   
 
### Histogram
    
```{r}
```
Per Capita Rates
===================================== 


About this Site
===================================== 

#### Last Update  

`r format(Sys.time(), "%d %B, %Y")`

#### Background
*"Nothing in life is to be feared, it is only to be understood. Now is the time to understand more, so that we may fear less" - Marie Curie*  


#### Code
Code and input data used to generate this web app are available on [Github](https://github.com/Erebus54/covid-geostats).

#### Sources 

**Status of COVID-19 cases in Ontario** : [Open Data Ontario](https://data.ontario.ca/dataset/status-of-covid-19-cases-in-ontario), updated daily, which presents a breakdown by region, case status, age range of cases 

**Census Profile, 2016 Census** : [StatsCan](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/index.cfm?Lang=E&TABID=1), queried this to obtain the areal size and population figures to compute the per capita metrics 

**Ministry of Health Public Health Unit Boundaries** : [Ontario GeoHub](https://geohub.lio.gov.on.ca/datasets/d8fba69152e4408dabfe70e85a2688d2_44), contains the geographic boundaries of all 34 Public Health Units (PHUs) in Ontario, based on Statistics Canada’s Health Region Boundary File (2018)

used to pull total population by health region and areal statistics to calculate per capita/density case loads 

#### Author 

Patrick Schnurbusch 

#### Contact 

patrick.schnurbusch@gmail.com
